
ADD SOME META DATA INFO 

SST NASA, ETC


## Reading and Loading the SST Data

Load the libraries 
```{r}
library(httr)
library(ncdf4)
library(tmap)
library(raster)
library(tictoc)
library(terra)
library(dplyr)
library(stringr)
library(lubridate)
library(ggplot2)
library(gganimate)

```


Make a mask
```{r}
# Make an empty raster of 0.008 degree resolution
x <- raster() # make an empty raster

crs(x) <- CRS("+proj=longlat +datum=WGS84") # Set the CRS
extent(x) <- extent(-123.32, -115.7, 30.59, 36.08) # set with the new extent
res(x) <- c(0.008, 0.008) # set resolution
values(x) <- 1:ncell(x) # fill in values 1-10873 by row


```

Read one image for testing
```{r}
file <- nc_open("/capstone/kelpgeomod/raw_data/SST/GRHSST_lvl4/20230101090000-JPL-L4_GHRSST-SSTfnd-MUR-GLOB-v02.0-fv04.1.dap.nc4")
sst <- ncvar_get(file, "analysed_sst")
sst_rast <- raster(sst)
plot(sst_rast)


# Flip the raster by 90 degrees to the left
flipped_data <- t(flip(sst_rast, 1))

plot(flipped_data)

```

Check the image info and use the mask
```{r}

# Set the extent of the raster to match the mask extent
raster_data <- setExtent(flipped_data, extent(x))

# Resample the raster to match the mask resolution
resampled_data <- resample(raster_data, x, method="bilinear")

# Apply the mask to the resampled raster
sst_masked <- mask(resampled_data, x)

#Check dimentions
dim(sst_masked)

plot(sst_masked)

```


# Creating a For Loop for the raster stack


Set up the directoty path where the files are located:
```{r}
data_dir <- "/capstone/kelpgeomod/raw_data/SST/GRHSST_lvl4"
```

Create a list of all the files in the directory:
```{r}
file_list <- list.files(data_dir, full.names=TRUE)
```

Crate a Data Frame that contains the information from the files names and creates a new column in a data frame:
```{r}
#Filter for the desired dates
start_date <- as.Date("2014-01-01")
end_date <- as.Date("2022-12-31")

file_df <- data.frame(file_name=file_list) %>%
  mutate(date=as.Date(str_extract(file_name, "\\d{8}"), format="%Y%m%d")) %>% 
  filter(date >= start_date & date <= end_date)

```

Create a function that reads in each file and convert it to a raster:

```{r}

ncdf_to_raster <- function(nc_file) {
  file <- nc_open(nc_file)
  sst <- ncvar_get(file, "analysed_sst")
  sst_rast <- raster(sst)
  
  # Flip the raster by 90 degrees to the left
  flipped_data <- t(flip(sst_rast, 1))
  
  # Set the extent of the raster to match the mask extent
  raster_data <- setExtent(flipped_data, extent(x))
  
  # Resample the raster to match the mask resolution
  resampled_data <- resample(raster_data, x, method="bilinear")
  
  # Apply the mask to the resampled raster
  sst_masked <- mask(resampled_data, x)
  
  nc_close(file) # Close the netCDF file
  
  return(sst_masked)
}

```

Do a test for January 2014 
```{r}

# Subset file_df to include only files from January 2014
jan_2014_files <- file_df %>% 
  filter(date < '2014-02-01')

# Use lapply() to create a list of rasters for January 2014
raster_list <- lapply(jan_2014_files$file_name, ncdf_to_raster)

# Convert to RasterStack object
jan_raster_stack <- stack(raster_list)

# Calculate the mean of all layers
jan_raster_mean <- calc(jan_raster_stack, mean)

plot(jan_raster_mean)

```


## Create a function that creates: 
a) A Data Frame with the year and month we are interested on
b) Apply the ncdf_to_raster function that reads in the image and converts it to a raster file
c) Stack all images from one month and executes the mean per month
```{r}

create_raster_stack <- function(year, month) {
  
  # Subset file_df to include only files from the given year and month
  files <- file_df %>% 
    filter(year(date) == year, month(date) == month)
  
  # Use lapply() to create a list of rasters for the given year and month
  raster_list <- lapply(files$file_name, ncdf_to_raster)
  
  # Convert to RasterStack object
  raster_stack <- stack(raster_list)
  
  # Calculate the mean of all layers
  raster_mean <- calc(raster_stack, mean)
  
  return(raster_mean)
  
}


```


```{r}

tic()
create_raster_stack(2017, 8)
toc()

```

## Create a bi for loop that:
a) Uses the function `create_raster_stack` to build a raster mean per month
b) Appends that new raster to a list 
c) Stack all rasters into a brick

```{r}
# Create empty list to store raster stacks
raster_stack_list <- list()

tic()
# Loop through all months from January 2014 to December 2022
for (year in 2014:2022) {
  for (month in 1:12) {
    
    # Create raster stack for current year and month
    raster_stack <- create_raster_stack(year, month)
    
    # Add raster stack to list
    raster_stack_list[[length(raster_stack_list) + 1]] <- raster_stack
    
  }
}

toc()

# Convert list of raster stacks to RasterBrick object
raster_brick <- brick(raster_stack_list)

class(raster_brick)

```

Create the monthly names for each raster monthly mean
```{r}
# Get current names of layers in raster_brick
current_names <- names(raster_brick)

# Create vector of new names in desired format
new_names <- paste0(month.abb, "-", rep(2014:2022, each=12))

# Rename layers in raster_brick with new names
names(raster_brick) <- new_names

```


Create a gif
```{r}
library(animation)

# Function to create the plot for each frame
plot_raster <- function(x) {
  plot(x, main = "", col = colorRampPalette(c("blue", "white", "red"))(100), 
       xlim = extent(-180, 180, -90, 90))
}

# Create the GIF
saveGIF({
  for (i in 1:nlayers(raster_brick)) {
    plot_raster(raster_brick[[i]])
    title(main = format(paste0(names(raster_brick)[i])), 
          font.main = 1, cex.main = 1.5)
  }
}, movie.name = "raster_brick_monthly.gif", interval = 0.5, ani.width = 800, ani.height = 400)


```


Quarterly base
```{r}

create_raster_stack_quarterly <- function(year, quarter) {
  
  # Define the start and end months for the given quarter
  start_month <- (quarter - 1) * 3 + 1
  end_month <- quarter * 3
  
  # Subset file_df to include only files from the given year and quarter
  files <- file_df %>% 
    filter(year(date) == year & month(date) >= start_month & month(date) <= end_month)
  
  # Use lapply() to create a list of rasters for the given year and quarter
  raster_list <- lapply(files$file_name, ncdf_to_raster)
  
  # Convert to RasterStack object
  raster_stack <- stack(raster_list)
  
  # Calculate the mean of all layers
  raster_mean <- calc(raster_stack, mean)
  
  return(raster_mean)
  
}
```

```{r}
create_raster_stack_quarterly(2015, 2)
```


```{r}
# Create empty list to store raster stacks
raster_quarter_stack_list <- list()

tic()
# Loop through all quarters from Q1 2014 to Q4 2022
for (year in 2014:2022) {
  for (quarter in 1:4) {
    
    # Create raster stack for current year and quarter
    raster_stack <- create_raster_stack_quarterly(year, quarter)
    
    # Add raster stack to list
    raster_quarter_stack_list[[length(raster_stack_list) + 1]] <- raster_stack
    
  }
}

toc()

# Convert list of raster stacks to RasterBrick object
raster_quarter_brick <- brick(raster_quarter_stack_list)

```


Save the raster bricks
```{r}

# Write RasterBrick Quarter object to file
writeRaster(raster_quarter_brick, filename = "/capstone/kelpgeomod/intermediate_data/sst/sst_quarter_raster", format = "GTiff")


# Write RasterBrick Monthly object to file
writeRaster(raster_brick, filename = "/capstone/kelpgeomod/intermediate_data/sst/sst_monthly_raster", format = "GTiff")

```




