---
title: "reclassify_usgs"
author: "Jessica French"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The USGS files have better coverage along the coast but have a more complicated classification scheme compared the the rasters from the California mapping project. In this notebook I am going to try and see what classifications in the USGS shape files overlap with the rocky areas in the shape files in order to reclassify.

```{r label = "read in libraries"}
library(terra)
library(tmap)
library(sf)
library(tidyverse)
```

```{r label = "set working directory"}
setwd("/Users/jfrench/Documents/MEDS/Capstone/DATA/Benthic_Habitat_2")
```

Lets start small scale, the sbc f g an h files overlap with carpenteria, ventura, santa barbara offshore files from the usgs. 

Start by reading in both sets of files

```{r label = "read in sbc f, g, h}
sbc_f_2msub <- rast("Habitat_tiffs/sbc_f_2msub.tif")
sbc_g_2msub <- rast("Habitat_tiffs/sbc_g_2msub.tif")
sbc_h_2msub <- rast("Habitat_tiffs/sbc_h_2msub.tif")
```

```{r label = "read in carpenteria, ventura, santa barbara"}
carpenteria <- st_read("Santa Barbara Channel _coastline/Carpinteria_Loon/Habitat_OffshoreCarpinteria/Habitat_OffshoreCarpinteria.shp") |> st_make_valid() |> 
  filter(Ind_ID == "h")
ventura <- st_read("Habitat_OffshoreVentura/Habitat_OffshoreVentura.shp") |> 
  st_make_valid() |> 
  filter(Ind_ID == "h")
santa_barbara <- st_read("Habitat_OffshoreSantaBarbara/Habitat_OffshoreSantaBarbara.shp") |>
  st_make_valid() |> 
  filter(Ind_ID == "h")
```

```{r label = "tmap"}
tmap_mode(mode = "view")

tm_shape(ventura) +
  tm_polygons(col = "Hab_Type", legend.show = FALSE) +
  tm_shape(carpenteria) +
  tm_polygons(col = "Hab_Type", legend.show = FALSE) +
  tm_shape(santa_barbara) +
  tm_polygons(col = "Hab_Type")
```

Whelp, as it turns out there is a column that simplifies the substrate to hard, soft, or mixed which matches two other files we have. 